---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ardglad

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of ardglad is to provide access to the Global Land Analysis & Discovery
(GLAD) Landsat Analysis Ready Data (ARD). The 
[GLAD ARD](https://www.glad.umd.edu/ard/home#format) represents a 16-day time 
series of globally consistent, tiled Landsat normalized surface reflectance 
from 1997 to the present operationally updated every 16 days.

This R package interacts with the [GLAD ARD API](https://www.glad.umd.edu/ard/home#download)
to download and process the data. 

## Installation

You can install the development version of ardglad like so:

``` r
# install.packages("pak")
pak::pkg_install("TESS-Laboratory/ardglad")
```

## Example

Here is an example of how to generate the urls for the GLAD ARD tiles for an 
example location in Rondônia, Brazil. 

```{r get-urls}
library(ardglad)

rondonia_eg <- sf::st_point(c(-62.625, -8.989)) |>
  sf::st_sfc(crs = 4326) |>
  sf::st_buffer(20000)

ron_urls <- ard_glad_urls(
  rondonia_eg,
  start_date = "2023-07-01",
  end_date = "2023-07-30"
)

print(ron_urls)
```

And now let's download the data and plot it. Note that although an object of 
class `ard_glad` is returned, in reality this is just a list of lists that 
contain information about each raster; to convert this to a more general use 
object for R, use `as.list` which will coerce the object to a list of 
`SpatRasters`. 

Note we can use `future` to parallelise some of the operations such as downloads 
and masking - by setting the plan here parallisation will be carried out 
automatically.

```{r download-data}
future::plan("multisession", workers = 3)

ron_ras <- ard_glad_download(ron_urls, "rondonia_glad")

print(ron_ras)

par(mfrow = c(1, 3))
plot(ron_ras)
```

We can also query specific moments through time, by specifying multiple 
`start_date` values, like so: 

```{r multi-date}
ron_urls_97_23 <- ard_glad_urls(
  rondonia_eg,
  c("1997-06-30", "2010-06-01", "2023-06-30")
)
```
Now we can download these data and mask clouds, shadows and haze, ready for 
further analysis:

```{r cloud-mask}
ron_ras_97_23_masked <- ard_glad_download(ron_urls_97_23, "rondonia_glad") |>
  ard_glad_mask()

par(mfrow = c(1, 3))
plot(ron_ras_97_23_masked)
```

As a bonus, here is a false colour composite:

```{r fcc}
par(mfrow = c(1, 3))
plot(ron_ras_97_23_masked, r = 4, g = 3, b = 2)
```