---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ardglad

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of ardglad is to provide access to the Global Land Analysis & Discovery
(GLAD) Landsat Analysis Ready Data (ARD). The 
[GLAD ARD](https://www.glad.umd.edu/ard/home#format) represents a 16-day time 
series of globally consistent, tiled Landsat normalized surface reflectance 
from 1997 to the present operationally updated every 16 days.

This package interacts with the [GLAD ARD API](https://www.glad.umd.edu/ard/home#download)
to download and process the data. The package is currently in development and
only provides functionality for returning the tile URLs.

## Installation

You can install the development version of ardglad like so:


``` r
# install.packages("pak")
pak::pkg_install("TESS-Laboratory/ardglad")
```

## Example

Here is an example of how to generate the urls for the GLAD ARD tiles for an 
example location in Rondonia, Brazil.

```{r example}
library(ardglad)

rondonia_eg <- sf::st_point(c(-61.604, -12.864)) |>
  sf::st_sfc(crs = 4326) |>
  sf::st_sf() |>
  sf::st_buffer(50000)

ard_glad_urls(rondonia_eg, "2023-01-01", "2023-03-30")
```

