---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ardglad

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of ardglad is to provide access to the Global Land Analysis & Discovery
(GLAD) Landsat Analysis Ready Data (ARD). The 
[GLAD ARD](https://www.glad.umd.edu/ard/home#format) represents a 16-day time 
series of globally consistent, tiled Landsat normalized surface reflectance 
from 1997 to the present operationally updated every 16 days.

This package interacts with the [GLAD ARD API](https://www.glad.umd.edu/ard/home#download)
to download and process the data. **The package is currently in development and
only provides functionality for returning the tile URLs**.

## Installation

You can install the development version of ardglad like so:


``` r
# install.packages("pak")
pak::pkg_install("TESS-Laboratory/ardglad")
```

## Example

Here is an example of how to generate the urls for the GLAD ARD tiles for an 
example location in Rondônia, Brazil.

```{r get-urls}
library(ardglad)

rondonia_eg <- sf::st_point(c(-61.604, -12.864)) |>
  sf::st_sfc(crs = 4326) |>
  sf::st_sf() |>
  sf::st_buffer(20000)

rondonia_urls <- ard_glad_urls(rondonia_eg, "2023-07-01", "2023-07-30")

print(rondonia_urls)
```

And now let's download the data and plot it.
```{r download-data}
rg <- ard_glad_download(rondonia_urls, "rondonia_glad")

# plot the result
par(mfrow = c(1, 3))
p <- mapply(
  \(x, y) {
    terra::plotRGB(x,
      r = 3, g = 2, b = 1,
      zlim = c(0, 7000),
      zcol = TRUE,
      stretch = "lin",
      axes = TRUE,
      mar = c(2, 2, 2, 1),
      main = y
    )
  },
  rg,
  names(rg)
)
```

